---
title: "Looking for CDI Bias: WG Comprehension"
author: "Nathan, George, and Mike"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = FALSE)

library(wordbankr)
library(tidyverse)
library(ggrepel)
library(mirt) 
library(here)
library(kableExtra)
library(gridExtra)
library(printr)
library(GGally)
library(ggpubr)
#library(parallel)
#library(doParallel)
#library(mclust)
#library(Rtsne)
#library(factoextra)
```

First, let's get some wordbank data for American English. We'll start with production data from the CDI: Words & Gestures (WG) form.
We'll look for DIF based on 1) sex and 2) SES (high vs. low). 
Questions: How do we choose a dozen anchor items (that we expect to be unbiased)?

```{r get-wordbank-data, eval=F, echo=F}
# Run Once
#get_instruments()

d_demo <- 
  get_administration_data(language = "English (American)", form = "WG") #%>%
  #bind_rows(get_administration_data(language = "English (American)", form = "WS"))

# 8-18 month-olds
table(d_demo$age)

#sum(is.na(d_demo$mom_ed)) # many mom_ed missing

```

Let's do comprehension data.

```{r get-wordbank-items, eval=F, echo=F}
d_long_wg <- get_instrument_data(language="English (American)", form = "WG") %>%
  mutate(understands = as.numeric(value == "yes" | value == "produces" | value == "understands")) # we'll start with production
# yes = understands; no = does not understand/produce, produces = produces (and understands)

en_wg <- get_item_data(language="English (American)", form="WG") %>%
  filter(type=="word") %>% select(-complexity_category, -item_id)

# subjects not producing any words yet (can't fit model with them)
bad_Ss = d_demo[which(d_demo$comprehension==0),]$data_id # 8 bad comprehension Ss

d_long_wg <- d_long_wg %>% left_join(en_wg %>% select(num_item_id, definition)) %>%
  filter(!is.na(definition))

d_mat <- d_long_wg %>% select(data_id, definition, understands) %>% 
  pivot_wider(id_cols = data_id, names_from = definition, values_from = understands)

d_long_wg <- d_long_wg %>% select(-value, -num_item_id)

save(d_mat, d_demo, en_wg, d_long_wg, file=here("data/en_wg_comp.Rdata"))
```

```{r load-data}
# load saved CDI:WG data
load(here("data/en_wg_comp.Rdata")) # d_demo, d_mat, en_wg (items)

d_demo <- d_demo %>% 
  filter(comprehension!=0, !is.na(sex)) %>% # can't fit children not producing words, or with NA sex in group model
  arrange(data_id)

# there are more IDs in d_mat than in d_demo; remove those
d_mat <- d_mat %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)

sids = d_mat$data_id
d_mat$data_id = NULL
d_mat = data.matrix(d_mat)
row.names(d_mat) = sids

table(d_demo$mom_ed)
# high SES = mom_ed >= Some College (739 high, 324 low-SES)
d_demo <- d_demo %>% mutate(ses_group = ifelse(as.numeric(mom_ed)>4, "high", "low"))

table(en_wg$category) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We will fit comprehension data from `r nrow(d_demo)` children for the WG form.

## Fit baseline 2PL model

```{r, eval=F}
m1 <- mirt(d_mat, 1, "2PL", 
           technical=list(NCYCLES=1000))

coefs_2pl <- as_tibble(coef(m1, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(m1, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat), 
                             ability = fscores(m1, method = "MAP")[,1])

sex_group = as.character(d_demo$sex)

save(file = here("data/eng_wg_comp_mod_2pl.Rds"), "m1", "coefs_2pl", "fscores_2pl") # ,
```

## GLIMMER Plots

```{r fit-rasch-group-models, echo=F, eval=F}
# first fit Rasch model to each group, and then compare resulting parameters

m1pl_m <- mirt(d_mat[which(d_demo$sex=="Male"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)
m1pl_f <- mirt(d_mat[which(d_demo$sex=="Female"),], 1, "Rasch", SE=T, SE.type='Oakes', verbose=T)

m1pl_h <- mirt(d_mat[which(d_demo$ses_group=="high"),], 1, "Rasch", SE.type='Oakes', verbose=T)
m1pl_l <- mirt(d_mat[which(d_demo$ses_group=="low"),], 1, "Rasch", SE.type='Oakes', verbose=T)

save(file = here("data/eng_wg_rasch_group_models.Rds"),
     "m1pl_m", "m1pl_f", "m1pl_h", "m1pl_l")

res_m = residuals(m1pl_m)
#m1pl_m_ci = boot.mirt(m1pl_m)
#itemplot(m1pl_m, 1, CE=T)
```


```{r, echo=F}
load(here("data/eng_wg_rasch_group_models.Rds"))

# coef(m1pl_m) shows CI_2.5 and CI_97.5 for d...iterate over and add to dataframes?
#ci_lo = coef(m1pl_m)$`baa baa`['CI_2.5','d']
#ci_hi = coef(m1pl_m)$`baa baa`['CI_97.5','d']

par_m = as_tibble(coef(m1pl_m, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_m, simplify = TRUE)$items))
par_f = as_tibble(coef(m1pl_f, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_f, simplify = TRUE)$items))

par_h = as_tibble(coef(m1pl_h, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_h, simplify = TRUE)$items))
par_l = as_tibble(coef(m1pl_l, simplify = TRUE)$items) %>%
      mutate(definition = rownames(coef(m1pl_l, simplify = TRUE)$items))

rasch_pars = tibble(definition = par_m$definition,
                    d_Male = par_m$d,
                    d_Female = par_f$d,
                    d_LowSES = par_l$d,
                    d_HighSES = par_h$d) %>%
  mutate(d_sex_diff = abs(d_Female - d_Male),
         d_ses_diff = abs(d_HighSES - d_LowSES))

sex_thresh = median(rasch_pars$d_sex_diff) + sd(rasch_pars$d_sex_diff)
ses_thresh = median(rasch_pars$d_ses_diff) + sd(rasch_pars$d_ses_diff)

p1 <- ggplot(rasch_pars, 
       aes(x = d_Male, y = d_Female)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_sex_diff > sex_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')

p2 <- ggplot(rasch_pars, 
       aes(x = d_LowSES, y = d_HighSES)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(rasch_pars, d_ses_diff > ses_thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Low SES") + 
  ylab("Item Ease for High SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

```{r, echo=F, fig.width=10, fig.height=5}
ggarrange(p1, p2)
```



# DIF Model: Sex

```{r, eval=F}
# https://rdrr.io/cran/mirt/man/DIF.html
# https://rdrr.io/cran/mirt/man/multipleGroup.html

sex_group = as.character(d_demo$sex)


# won't work if some items are all 0 (or 1?) for one sex
mg1 <- multipleGroup(d_mat, 1, group = sex_group,
                     technical=list(NCYCLES=4000)) 
#mg1 <- multipleGroup(d_mat, 1, group = sex_group, 
#                     invariance = c("free_var", "free_means")) 

mg2 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))

# constrain slopes, freely estimate latent group means (reference group constrained to a vector of 0)
mg3 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes', 'free_means'), verbose = T)

# constrain slopes, allow varying group means and variances -- very slow!
mg4 <- multipleGroup(d_mat, model = 1, group = sex_group, technical=list(NCYCLES=4000), 
                     invariance =c('slopes','free_means', 'free_var'), verbose = T)

# fixed means, varying group variance
mod_sc1 <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts', 'free_var'))
# fully-constrained
mod_fullconstrain <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

# constrain intercepts to be equal across groups, but allow group slopes to vary
m_int <- multipleGroup(d_mat, 1, group = sex_group, technical=list(NCYCLES=2000),
                     invariance = 'intercepts')

anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
anova(mg3, mg4) # AIC/BIC prefer mg3, with varying group means but equal variance

anova(mod_sc1, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc1 by AIC
anova(m_sc2, mod_fullconstrain) # mod_fullconstrain preferred by BIC, sc2 by AIC

anova(mod_fullconstrain, mg2) # AIC and BIC both prefer slope-invariant model

anova(m_int, mod_fullconstrain) # AIC prefers intercept-constrained model; BIC prefers fully-constrained


coef(mg3)[[2]]$GroupPars # Female intercept = 0.047 (Male = 0) 

#extract.group(m_sc2, "Male")
#extract.group(m_sc2, "Female")

residuals(mod_sc1)

save(file = here("data/eng_wg_comp_sex_sexDIF.Rds"), 
     mg1, mg2, mg3, mg4, m_sc2, mod_sc1, mod_fullconstrain) 
```

```{r}
load(here("data/eng_wg_comp_sex_sexDIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

extract_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$Male$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Male$items)) %>%
    select(-g, -u) %>% 
    rename(d_m = d)
  Fit = as_tibble(coef(group_model, simplify=T)$Female$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$Female$items)) %>%
    select(-g, -u) %>%
    rename(d_f = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_f - d_m,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}


# slope-invariant, varying difficulty:
mg2_df = extract_group_df(mg2)
mg3_df = extract_group_df(mg3)

# same item estimates with/without free_means? 
cor(mg2_df$d_f, mg3_df$d_f) # .999
cor(mg2_df$d_m, mg3_df$d_m) # .999


anova(mg1, mg2) # AIC and BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer mg3, with varying group means
```

### Items with a large difference in difficulty across sex

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg3_df$d_diff_abs) + sd(mg3_df$d_diff_abs)
big_dif = mg3_df[which(mg3_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
#fem = mg2s[[1]]$rotF
#mal = mg2s[[2]]$rotF
big_dif %>% relocate(definition) %>%
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```


### Plot item ease for males vs. females

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg3_df, 
       aes(x = d_m, y = d_f)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for Males") + 
  ylab("Item Ease for Females") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant free group means model (mg3).

```{r}
sex_anchors = mg3_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_m, .after=d_f)

sex_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```


## Ability distribution comparison

What are the ability distributions by sex in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=sex)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg3 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg3, method = "MAP")[,1])

fscores_mg2 <- tibble(data_id = rownames(d_mat), 
                             ability_sex = fscores(mg2, method = "MAP")[,1])

# very similar
#plot(fscores_mg3$ability_sex, fscores_mg2$ability_sex, pch='.')

d_demo <- d_demo %>% left_join(fscores_mg3 %>% 
                                 mutate(data_id = as.numeric(data_id))) 

ggpairs(d_demo, columns = c("age","comprehension","ability","ability_sex"),
        ggplot2::aes(colour=sex, alpha=.5)) + theme_classic()
```

```{r}
d_demo %>% group_by(sex) %>% 
  summarise(ability_2pl = mean(ability),
            ability_sex = mean(ability_sex)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```

Average male ability is slightly higher in the sex group model (i.e., there is a larger female advantage in the standard 2PL model), showing that the 36 items that are easier for females boost female ability estimates.


## SES-related DIF

Now we will look for SES-based DIF, using mother's education as a proxy for SES. 
Here's the distribution of mom_ed in our data:

```{r, mom-ed-tab, echo=F}
table(d_demo$mom_ed) %>% kable(format = "html", table.attr = "style='width:30%;'")
```

We (arbitrarily) define high SES as their mother having at least some college (N=`r length(which(d_demo$ses_group=="high"))`), and low SES as having at most secondary education (N=`r length(which(d_demo$ses_group=="low"))`).


```{r, eval=F}
ses_group = as.character(d_demo$ses_group)

mg1 <- multipleGroup(d_mat, 1, group = ses_group, SE = T, technical=list(NCYCLES=4000)) 
mg2 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = 'slopes', verbose = T)
mg3 <- multipleGroup(d_mat, model = 1, group = ses_group, technical=list(NCYCLES=4000), 
                     invariance = c('slopes', 'free_means'), verbose = T)

anova(mg1, mg2) # AIC/BIC prefer slope-invariant model (mg2)
anova(mg2, mg3) # AIC/BIC prefer slope-invariant model without varying group mean

#DIF(mg2, c('d')) # No hyper-parameters were estimated in the DIF model. For effective DIF testing, freeing the focal group hyper-parameters is recommended.

m_sc2 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                       invariance=c('slopes', 'intercepts', 'free_var','free_means'))
mod_sc1 <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000), #fixed means
                             invariance=c('slopes', 'intercepts', 'free_var'))
mod_fullconstrain <- multipleGroup(d_mat, 1, group = ses_group, technical=list(NCYCLES=2000),
                             invariance=c('slopes', 'intercepts'))

#extract.group(m_sc2, "high")
#extract.group(m_sc2, "low")

residuals(mod_sc1)

save(file = here("data/eng_wg_comp_SES_DIF.Rds"), 
     mg1, mg2, mg3, m_sc2, mod_sc1, mod_fullconstrain)
```

```{r}
load(here("data/eng_wg_comp_SES_DIF.Rds"))
load(here("data/eng_wg_comp_mod_2pl.Rds"))

extract_ses_group_df <- function(group_model) {
  Mit = as_tibble(coef(group_model, simplify=T)$high$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$high$items)) %>%
    select(-g, -u) %>% 
    rename(d_hi = d)
  Fit = as_tibble(coef(group_model, simplify=T)$low$items) %>%
    mutate(definition = rownames(coef(group_model, simplify=T)$low$items)) %>%
    select(-g, -u) %>%
    rename(d_lo = d)

  combo <- Mit %>% left_join(Fit) %>%
    mutate(d_diff = d_hi - d_lo,
           d_diff_abs = abs(d_diff)) 
  return(combo)
}

mg2_df = extract_ses_group_df(mg2)
mg3_df = extract_ses_group_df(mg3)

coef(mg3)[[2]]$GroupPars # low-SES group (high is ref) intercept = .25
```



### Items with a large difference in difficulty across SES

```{r}
# items showing > median + 1SD of absolute difference in difficulty have a lot of DIF
thresh = median(mg2_df$d_diff_abs) + sd(mg2_df$d_diff_abs)
big_dif = mg2_df[which(mg2_df$d_diff_abs > thresh),]

#mg2s <- summary(mg2)
big_dif %>% relocate(definition) %>% 
  arrange(desc(d_diff_abs)) %>% kable(digits = 2, format = "html", table.attr = "style='width:40%;'")
```


### Plot item ease for high vs. low SES

We label the `r nrow(big_dif)` items with absolute ease difference of more than median + 1SD = `r round(thresh,2)`.

```{r, fig.width=8, fig.height=6}
ggplot(mg2_df, 
       aes(x = d_lo, y = d_hi)) + 
  geom_point(alpha=.7) + 
  ggrepel::geom_label_repel(data = filter(mg2_df, d_diff_abs > thresh), 
                            aes(label = definition), max.overlaps = 20) + theme_bw() +
  xlab("Item Ease for low-SES") + 
  ylab("Item Ease for high-SES") + 
  geom_abline(intercept=0, slope=1, linetype='dashed')
```

### Anchor items

For our DIF analysis, let's pick 10 anchor items that show the smallest difficulty difference in the slope-invariant model (mg2).

```{r ses-anchors, echo=F}
ses_anchors = mg2_df %>% arrange(d_diff_abs) %>% head(10) %>% 
  relocate(a1, .after=definition) %>%
  relocate(d_hi, .after=a1)

ses_anchors %>% kable(digits = 3, format = "html", table.attr = "style='width:40%;'")
```


## Ability distribution comparison

What are the ability distributions by SES in the 2PL model?

```{r}
d_demo <- d_demo %>% left_join(fscores_2pl %>% mutate(data_id = as.numeric(data_id)))

d_demo %>% ggplot(aes(x=ability, fill=ses_group)) + 
  geom_density(alpha=.3) + 
  theme_classic()
```

```{r}
# get ability scores from sex group model with varying difficulty
fscores_mg2 <- tibble(data_id = rownames(d_mat[which(!is.na(d_demo$ses_group)),]), 
                             ability_ses = fscores(mg2, method = "MAP")[,1])

d_demo_ses <- d_demo %>% 
  filter(!is.na(ses_group)) %>% 
  left_join(fscores_mg2 %>% mutate(data_id = as.numeric(data_id)))

ggpairs(d_demo_ses, columns = c("age","comprehension","ability","ability_ses"),
        ggplot2::aes(colour=ses_group, alpha=.5)) + theme_classic()
```

```{r, echo=F}
d_demo_ses %>% group_by(ses_group) %>% 
  summarise(ability_2pl = mean(ability),
            ability_ses = mean(ability_ses)) %>%
  kable(digits = 2, format = "html", table.attr = "style='width:30%;'")
```
